Symmetric Circuits for Rank Gates

### 2017-07-26 sections/introduction.tex
26/07/2017

\documentclass[12pt]{report}

% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry} % set the margins to 1in on all sides
\usepackage{graphicx} % to include figures
\usepackage{amsmath} % great math stuff
\usepackage{amsfonts} % for blackboard bold, etc
\usepackage{amsthm} % better theorem environments
\usepackage{xspace}

% various theorems, numbered by section

\newtheorem{thm}{Theorem} \newtheorem{claim}[thm]{Claim}
\newtheorem{remark}[thm]{Remark} \newtheorem{definition}[thm]{Definition}
\newtheorem{lem}[thm]{Lemma} \newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary} \newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\reals}{\mathbb{R}}
\newcommand{\nats}{\mathbb{N}}

\newcommand{\logic}[1]{\text{\upshape #1}\xspace}
\newcommand{\FPC}{\logic{FPC}}
\newcommand{\FO}{\logic{FO}}
\newcommand{\FP}{\logic{FP}}
\newcommand{\FPR}{\logic{FPrk}}
\newcommand{\CPTC}{\logic{$\tilde{\text C}$PT(Card)}}

\newcommand{\PT}{\ensuremath{\mathrm{P}}\xspace}

\newcommand{\bd}[1]{\mathbf{#1}} % for bolding symbols
\newcommand{\RR}{\mathbb{R}} % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}} % for Integers
\newcommand{\MB}{\mathbb{B}_{\matsym}} % for Integers
\newcommand{\SB}{\mathbb{B}_{\sym}} % for Integers
\newcommand{\col}[1]{\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}
\newcommand{\stab}{\text{\textbf{Stab}}}
\newcommand{\setstab}{\text{\textbf{SetStab}}}
\newcommand{\matstab}{\text{\textbf{MatStab}}}
\newcommand{\inv}{\text{\textbf{inv}}}
\newcommand{\aut}{\text{\textbf{Aut}}}
\newcommand{\kernal}{\text{\textbf{ker}}}
\newcommand{\alt}{\text{\textbf{Alt}}}
\newcommand{\sym}{\text{\textbf{Sym}}}
\newcommand{\std}{\text{\textbf{std}}}
\newcommand{\rank}{\text{\textbf{rk}}}
\newcommand{\orb}{\text{\textbf{Orb}}}
\newcommand{\SP}{\text{\textbf{SP}}}
\newcommand{\SPs}{\text{\textbf{SPs}}}
\newcommand{\maj}{\text{\textbf{maj}}}
\newcommand{\dom}{\text{\textbf{Dom}}}
\newcommand{\img}{\text{\textbf{Img}}}
\newcommand{\child}{\text{child}}
\newcommand{\countgate}{\text{count}}
\newcommand{\agree}{\text{AGREE}}
\newcommand{\supp}{\text{SUPP}}
\newcommand{\rankparam}{\text{PARAM}}
\newcommand{\consp}{\text{sp}}
\newcommand{\consps}{\text{sps}}
\newcommand{\row}{\text{row}}
\newcommand{\column}{\text{col}}
\newcommand{\rowsp}{\text{r}}
\newcommand{\colsp}{\text{c}}
\newcommand{\EV}{\text{EV}}
\newcommand{\type}{\text{type}}
\newcommand{\types}{\text{types}}
\newcommand{\spstab}[1]{\stab (\consp (#1))}

\newcommand{\Alpha}{A}
\newcommand{\matsym}{\text{\textbf{MatSym}}}
\begin{document}

\title{Symmetric Circuits for Rank Gates}

\author{Anuj Dawar and Gregory Wilsenach \\ 
Computer Laboratory \\
University of Cambridge}

\maketitle

\tableofcontents

\chapter{Introduction}

A finite relational structure, such as a graph, can be represented as a binary
string in a standard way (for instance, by enumerating its adjacency matrix).
When we speak of the computational complexity of properties of graphs, we assume
such an encoding in order to present the structure as input to a machine or
circuit or other appropriate computational model. The properties of structures
we are interested in studying are invariant under isomorphisms. This requirement
translates into an invariance condition on sets of strings---say that a language
$L \subseteq \{0,1\}^\star$ is a \emph{graph property} if, whenever $x \in L$
for a string $x$ representing a graph $G$, we have $y\in L$ for any string $y$
obtained from $x$ by re-ordering the vertices of $G$. The central open question
in descriptive complexity---whether there is a logical characterisation of the
polynomial-time decidable properties of finite relational structures--- can then
be understood as the question of characterizing those languages $L$ that are in
$\PT$ and are graph properties.

@@

This motivates the study of symmetric circuit families that involve gates which
themselves may compute non-symmetric functions. In particular, motivated by the
study of rank logic~\cite{DGHL09,GP15}, we consider gates that are
\emph{matrix-symmetric}. These are gates computing a Boolean function $f:
\{0,1\}^{A \times B} \rightarrow \{0,1\}$ which is invariant under arbitrary
permutations of $A$ and $B$. A typical example is the function computing the
rank of an $A \times B$ $0$-$1$ matrix. We show that the support theorem of
Anderson and Dawar can be shown to hold, even in the presence of
matrix-symmetric gates. This allows us to show that definability in the logic
fixed-point with rank ($\FPR$) is the same as decidability by
polynomially-uniform families of symmetric circuits with rank gates.

% Another extension of $\FPC$ that has received much attention in the descriptive
% complexity literature is the class $\CPTC$ of \emph{choiceless polynomial time
%   with counting}. The work of~\cite{AD14} leaves open the question of obtaining
% a circuit characterization of this class and it was suggested there that this
% would require a relaxation of the notion of symmetry used in defining symmetric
% circuits. In particular, while $\CPTC$ is based on a machine model designed to
% preserve symmetry by forbidding choices, the polynomial time and space
% restrictions apply int he context of the symmetries fo a \emph{single input
%   structure}. In this paper, we consider a relaxation of the notion of symmetric
% circuit intended to capture exactly this intuition. However, we are able to show
% that this does not yield a characterization of $\CPTC$. In particular, we show
% that no family in this class of relaxed symmetric circuits can distinguish odd
% from even Cai-F\"urer-Immerman graphs, something we know is do-able in
% $\CPTC$~\cite{DRR08}.

\chapter{Background}
In this section we present basic background on logics, first-order structures,
and symmetric circuits.

\section{Logic and Structures}
A \emph{relational vocabulary} (usually denoted by $\tau$) is a finite sequence
of relation symbols $R_1, \ldots, R_k$, such that for all $i \in [k]$, $R_i$ has
arity of $r_i \in \nats$.

A $\tau$-structure $\mathcal{A}$ is a tuple $\langle U, R^{\mathcal{A}}_1 ,
\ldots , R^{\mathcal{A}}_k \rangle$, where $U$ is a set and called the
\emph{universe} of $\mathcal{A}$ and for all $i \in [k]$ the relations
$R^{\mathcal{A}}_i \subseteq U^{r_i}$. The elements of $U$ are called the
elements of $\mathcal{A}$. A \emph{multi-sorted} structure is one whose universe
is a disjoint union, where each set in the union is called a \emph{sort}. The
size of a structure is the cardinality of its universe.

All structures in this paper are finite.

\section{First Order, Fixed-Point, Counting}

Let $\FO(\tau)$ denote \emph{first order logic} with respect to the vocabulary
$\tau$. \emph{Fixed-Point Logic} is the extension of $\FO(\tau)$ with an
inflationary fixed-point operator (see \cite{} for details). For logic
$\mathcal{L}$, a formula $\phi \in \mathcal{L}(\tau)$, $x$ a $k$-length tuple of
variables, $\mathcal{A}$ a (finite) structure, and $a \in U^k$, we have that
$\phi (x)$ denotes that the variables in $x$ are free in $\phi$ and we write
$\mathcal{A} \models_{\mathcal{L}} \phi[a]$ iff the the formula $\phi$ is true
in $\mathcal{A}$ with respect to $\mathcal{L}$ with the variables $x$ assigned
to $a$. The context usually makes the logic obvious, and in that case we usually
drop the subscript.

We also recall the definition of \emph{fixed-point logic with counting} $\FPC$,
an extension $\FP$ with counting quantifiers. For a structure $\mathcal{A}$,
Formulas of this logic are evaluated over a structure $\mathcal{A}$ by extending
that structure to $\mathcal{A}^\leq$, a two sorted extension of $\mathcal{A}$ by
an ordered number sort $\{1, \ldots, \vert U\vert \}$. Variables are taken to be
either point variables, ranging over $U$ or number variables ranging over
$[\vert U \vert]$. The counting quantifiers are of the form $\#_x$, where $x$ is
either a point or number variable and for $\phi(x) \in \FPC(\tau)$, $\#_x
\phi(x)$ is a term denoting the element of the number sort corresponding to
$\vert \{a \in \mathcal{A} | \mathcal{A} \models \phi[a]\}\vert$. Note
additionally that $k$-tuples of number variables can be used to encode numbers
in $\FPC$ of size up to $\vert U\vert^k$. For more details on $\FPC$ please see
\cite{}.

\section{Rank Logic}
This still needs to be filled in

\section{Symmetric Circuits}
This still needs to be filled in

\chapter{Symmetric Circuits}
In this section we discuss symmetric circuits, their limitations and extensions
of this model.

\section{Symmetric Circuits and Limitations}
Symmetric circuits, as developed by Anderson and Dawar, are, importantly defined
over (complete) Boolean basis consisting of symmetric functions. The following
result shows that once a majority gate has been included in this basis any
symmetric function can be computed using polynomial-size symmetric circuits, and
as such no additional symmetric function added to the Basis improves the power
of the model.

Recall from Anderson and Dawar \cite{AndersonD17} we have that $\mathbb{B}_{\std} = \{
\neg , \wedge , \lor \}$ and $\mathbb{B}_\maj = \{ \maj \} \cup \mathbb{B}$.

Let $F: \{0,1\}^* \rightarrow \{0,1\}$ be a symmetric Boolean function. Since
$F$ is symmetric we note that for a fixed size input the output of $F$ is
entirely determined by the number of 1's in its input. Then let
$c_{F}:\mathbb{N} \rightarrow 2^{\mathbb{N}}$ define a function where $c_{F}(n)$
is the set of all $m \leq n$ such that for all $\vec{x} \in \{ 0,1 \}^n$ with
$m$ 1's we have $F (\vec{x}) = 1$. Clearly any symmetric Boolean function $F$ is
entirely determined by $c_{F}$.

\begin{prop}
  \label{prop:fuctions-maj}
  There is a polynomial $p(k)$ such that for any symmetric function $F$ and a
  given $k \in \mathbb{N}$ there is a circuit $C_k$ on $k$ inputs over the basis
  $\mathbb{B}_\maj$ which is symmetric, constant depth and with width bounded by
  $p(k)$.
\end{prop}

\begin{proof}
  We define the circuit $C_k$ for inputs $\vec{x} = ( x_1, \ldots, x_k )$.

  For $a \in \mathbb{N}$ we define a get $\countgate_a$ by
  \begin{align*}
    &\countgate_a = \maj (x_1, \ldots, x_k, \underbrace {0, \ldots, 0}_{2a -
      k}) \land \neg \maj (x_1, \ldots , x_k, \underbrace{0, \ldots,  0}_{2a - k + 2})\text{ if $a \geq \frac{k}{2}$,} \\
    &\countgate_a = \maj (x_1, \ldots, x_k, \underbrace {1, \ldots, 1}_{x -
      2a}) \land \neg \maj (x_1, \ldots , x_k, \underbrace{1, \ldots,
      1}_{k - 2a -2}) \text{ if $a < \frac{k}{2}$. }
  \end{align*}

  Then let $g = \bigvee_{a \in c_{F_i}(k)}\countgate_a$ and let $C_{k}$ be the
  circuit with input gates labeled by $\vec{x}$ and output gate $g$.

  t is easy to see that $C_k$ is constant depth and it's width is a polynomial
  in $k$. We have that in each layer of $C_{g'}$ each gate is connected to all
  gates in the previous layer, and as such the circuit is symmetric.
\end{proof}

The above proposition has a straight forward application to circuit
characterisations.

\begin{thm}
  Let $F = \{F_i : i \in I \}$ be a family of symmetric Boolean functions where
  $F_i: \{0,1\}^* \rightarrow \{ 0,1 \}$.

  Let $(C_n)_{n \in \mathbb{N}}$ be family of symmetric circuits over the
  Boolean basis $\mathbb{B} \cup F$, where $C_n$ is a circuit on structures of
  size $n$, and the size of each circuit in the family is bounded by some
  function $f(n)$. Then there exists a polynomial $q(n)$ and a family of
  symmetric circuits $(C_n')_{n \in \mathbb{N}}$ over $\mathbb{B}_\maj$, where
  $C_n'$ is a circuit on structures of size $n$ and $\vert C_n' \vert \leq
  q(f(n))$.
\end{thm}

\begin{proof}
  From $C_n$ we construct $C_n'$ in the obvious way. For each gate $g \in C_n$
  of type $F_i$ we have a symmetric circuit $C_g$ from Proposition
  \ref{prop:function-maj} that computes the same function as $g$. Then let
  $C_n'$ be $C_n$ but with each gate $g$ replaced by $C_g$. It is easy to see
  that $C_n'$ is symmetric. We also have that each gate $g$ must have at most
  $f(n)$ inputs, and the size of $C_g$ is bounded by $p(f(n))$. Thus the size of
  $C_n'$ is bounded by $f(n)p(f(n))$.
\end{proof}

\section{G-Symmetric Functions}
In order to extend the notion of a circuit by extending the basis we need to
consider functions that are not symmetric in the full sense of the term but have
some weakened notion of symmetry.

\begin{definition}
  Let $f: \{0,1\}^X \rightarrow \{0,1\}$ be a function. Let $G \subseteq \sym_X$
  be a subgroup. We say that $f$ is \emph{$G$-symmetric} if for all $\sigma \in
  G$, $x \in X$, $f(\sigma x) = f(x)$.
\end{definition}

We can also introduce similar notions in the case of a two dimensional Boolean
string.

\begin{definition}
  Let $f: \{0,1\}^{A \times B} \rightarrow \{0,1\}$ be a function. Let $G
  \subseteq \sym_{A \times B}$ be a subgroup. We say that $f$ is
  \emph{matrix-symmetric} if it is $\sym_A \times \sym_B$-symmetric and we say
  that $f$ is \emph{graph-symmetric} if $A = B$ and it is $\sym_A$-symmetric
  (with $\sym_A$ acting on both $A$ and $B$).
\end{definition}

So if we think of $f: \{0,1\}^{A \times B} \rightarrow \{0,1\}$ as representing
a matrix, then matrix-symmetry corresponds to the property of the function being
invariant under row-column permutations. If, instead, we have that $A = B$ and
we then think of $f$ as encoding the adjacency matrix of a graph, then
graph-symmetry corresponds to the function being constant on isomorphism
classes.

Importantly, many natural functions of interest are matrix symmetric. For
example, the function that computes the rank of the matrix over $\mathbb{F}_2$.
or a thresholded rank function, for example the rank of the matrix over
$\mathbb{F}_p$ being larger then $r$, for some particular $(p, r) \in
\mathbb{N}$.

\section{Symmetric Circuits with Matrix-Symmetric Gates}

In this section we develop the corresponding notion of a symmetric circuit, ones
which allow for the inclusion of matrix-symmetric gates. In particular, these
circuits are defined using two sets of Boolean functions. The first, denoted by
$\mathbb{B}_\sym$, is taken to be the symmetric Boolean basis (either
$\mathbb{B}_\std$ or $\mathbb{B}_\maj$). The second, denoted by
$\mathbb{B}_\matsym$ is taken to be the matrix-symmetric basis, consisting of a
collection of matrix-symmetric functions.

\begin{definition}[Circuits on Structures]
  For $\mathbb{B}_{\sym}$ a basis of Boolean symmetric functions,
  $\mathbb{B}_{\matsym}$ a basis of Boolean matrix-symmetric functions and
  $\tau$ a set of relation symbols, we define a $(\mathbb{B}_\sym,
  \mathbb{B}_\sym, \tau)$-circuit $C_n$ computing a $q$-ary query $Q$ is a
  structure $\langle G, W, \Omega, \Sigma, \Lambda, L\rangle$.
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G$ is called the set of gates of $C_n$ and $\vert C_n \vert := \vert G
    \vert$.
  \item $W \subseteq G \times G$, where $W$ is called the wires of the circuit.
    $(G,W)$ must be a directed acyclic graph. For $g \in G$ we $H_g := \{ h \in
    C_n : W(h,g)\}$ be the set of children of $g$.
  \item $\Omega$ is an injective function from $[n]^q$ to $G$. The gates in the
    image of $\Omega$ are called the output gates. When $q = 0$, $\Omega$ is a
    constant function mapping to a single output gate.
  \item $\Sigma$ is a function from $G$ to $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab \uplus \tau \uplus \{0,1\} $ which maps input gates to
    $\tau \uplus \{0,1\}$ and where $\Sigma^{-1} (0) \leq 1$ $\Sigma^{-1} (1)
    \leq 1$ and the internal gates get mapped into $\mathbb{B}_\sym \uplus
    \mathbb{B}_\matstab$. Gates mapped to $\tau$ are called relational gates and
    gates mapped to 1 or 0 are called constant gates.
  \item $\Lambda$ is a sequence of injective functions $(\Lambda_R)_{R \in
      \tau}$ where for each $R \in \tau$, $\Lambda_R$ maps each relational gate
    $g$ with $R = \Sigma (g)$ to the tuple $\Lambda_R (g) \in [n]^r$, where $r$
    is the arity of the symbol $R$. When no ambiguity arises we write $\Lambda
    (g)$ for $\Lambda_R (g)$.
  \item $L$ assigns labels to the gates (and their inputs) in $C_n$. Let $g \in
    G$ be an internal gate and $H$ be the gates input to $g$. Then $L(g) : A
    \times B \rightarrow H$, where $L(g)$ must be surjective and $A, B \subseteq
    [n]$. We call this the \emph{matrix labelling} of $g$.
  \end{itemize}
\end{definition}

Note the additional requirement that each gate have an associated labelling in
order to facilitate the evaluation of the gate.

Given some finite $\tau$-structure $\mathcal{A}$ of size $n$ and some bijection
$\gamma: U \rightarrow [n]$, the evaluation of some $(\mathbb{B}_\sym,
\mathbb{B}_\sym, \tau)$-circuit) circuit $C$ proceeds by recursively evaluating
gates. Here the evaluation of the gate $g$ is denoted by $C[\gamma
\mathcal{A}](g)$. If $g$ is anything but a matrix-symmetric gate, then the
evaluation is as per the symmetric circuits of Anderson and Dawar \cite{}. If
$g$ is matrix-symmetric then the evaluation is given by applying the Boolean
operation $\Sigma(g)$ to the matrix $L^{\gamma}:\dom(L(g)) \rightarrow \{0,1\}$
defined by $L^{\gamma}(i,j) = C[\gamma \mathcal{A}](L(g)(i,j))$.

\begin{definition}[Invariant Circuit\cite{AndersonD17}]
  Let $C_n$ be a $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, computing
  some $q$-ary query. We say $C_n$ is \emph{invariant} if for every
  $\tau$-structure $\mathcal{A}$ of size $n$, $a \in \mathcal{A}^q$, and
  bijections $\gamma_1, \gamma_2: U \rightarrow [n]$ we have that $C[\gamma_1
  \mathcal{A}](\Omega (\gamma_1 a)) = C[\gamma_2 \mathcal{A}](\Omega (\gamma_2
  a))$.
\end{definition}

\begin{definition}[Automorphism]
  let $C = \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$ be a
  $(\mathbb{B}_\sym, \mathbb{B}_\matsym, \tau)$-circuit computing at $q$-ary
  query on structures of size $n$. Let $\sigma \in \sym_n$ and $\pi: G
  \rightarrow G$ be a bijection such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item for all gates $g, h \in G$, $W(g,h)$ iff $W(\pi g, \pi h)$,
  \item for all output tuples $x \in [n]^q$, $\pi \Omega (x) = \Omega (\sigma
    x)$,
  \item for all gates $g \in G$, $\Sigma (g) = \Sigma (\pi g)$,
  \item for each relational gate $g \in G$, $\sigma \Lambda (g) = \Lambda (\pi
    g)$, and
  \item for each internal gate $g$ if $\Sigma (g) \in \mathbb(B)_\matsym$ then
    we have that $L(\pi g) \sim \pi \cdot L(g)$.
  \end{itemize}

  We call $\pi$ an \emph{automorphism} of $C$, and we say that $\sigma$
  \emph{induces the automorphism} $\pi$. The group of automorphisms of $C$ is
  called $\aut_n (C)$.
\end{definition}

\begin{definition}[Symmetric\cite{AndersonD17}]
  A circuit $C$ on structures of size $n$ is called \emph{symmetric} if every
  $\sigma \in \sym_n$ induces an automorphism on $C$.
\end{definition}

It follows for any symmetric circuit $C_n$ there is a mapping from $\sym_n$ to
$\aut_n(C)$, from $ \sigma$ to the induced automorphism. This homomorphism is
injective so long as a single element of $[n]$ appears in the in the label of
some input gate of $C$ (as then all elements appear by
symmetry)\cite{AndersonD17}. In this paper we always assume that there is always
one such element as otherwise all inputs are constant, and so the circuit just
computes a constant function. In order to assure this homomorphism is surjective
Anderson and Dawar \cite{AndersonD17} introduce the notion of a \emph{rigid}
circuit.

\begin{definition}[Rigidity]
  Let C be a $(\mathbb{B}_\sym, \mathbb{B}_\sym, \tau)$-circuit, where $C =
  \langle G, W, \Omega, \Sigma, \Lambda, L\rangle$. Say that $C$ is rigid if
  there are no internal gates $g, g' \in G$ such that $\Sigma(g) = \Sigma (g')$,
  $\Omega^{-1}(g) = \Omega^{-1}(g')$, and for every $g'' \in G$, $W(g'', g')$
  and $W(g,g'')$ iff $W(g', g'')$.
\end{definition}

Another property which simplifies our analysis is the property of having
\emph{bijective labels}.

\begin{definition}
  We say that a circuit $C$ has bijective labels if for each gate $g$ in $C$,
  $L(g)$ is a bijection.
\end{definition}

We prove in a later section that a circuit may be transformed in polynomial time
into an equivalent circuit that is both rigid and has bijective labelings. Hence
we may assume these two properties without a loss of generality.

With the assumption of rigidity in place, we abuse notation for permutations and
let $\sigma \in \sym_n$ also denote the induced automorphism.

We finally want to define symmetric circuits with rank gates.

\begin{definition}
  Let $\mathbb{B}_{\matsym}$ consist of functions of the form $f_{p,r}:
  \{0,1\}^{A \times B} \rightarrow \{0,1\}$ for $A, B$ initial segments of
  $\mathbb{N}$ and $f, p \in \mathbb{N}$, with $p$ prime, and such that for
  $f_{p,r}(M) = 1$ iff the matrix $M: A \times B \rightarrow \{0,1\}$ has rank
  at least $r$ over the field $\mathbb{F}_p$. A \emph{symmetric circuit with
    rank gates} is a $(\mathbb{B}_\maj, \mathbb{B}_\matsym, \tau)$-circuit, for some
  vocabulary $\tau$.
\end{definition}

The natural restriction to consider on families of circuits is uniformity.

\begin{definition}
  Let $(C_n)_{n \in \mathbb{N}}$ be a family of Boolean circuits. We say that
  $(C_n)_{n \in \mathbb{N}}$ is \emph{$P$-uniform} if the mapping $n \mapsto C_n$ is
  computable in polynomial time.
\end{definition}

We are now ready to state what will be the main theorem of this paper.

\begin{thm}[Main Theorem]
 A graph property is decidable by a $P$-uniform family of symmetric circuits
 with rank gates if, and only if, it is definable by an FPR sentence.
\end{thm}

% \section{Circuits to FPR}

% In this section I discuss how to determine the rank a matrix labelling at some
% gate $g$. This is a rough copy written for the purpose of discussion. In this
% section we fix some circuit $C_n$ and some rank gate $g$ in $C_n$ with a set
% of children $H$ and matrix labelling $L: [a] \times [b] \rightarrow H$ (which
% we assume WLOG is a bijection).

\chapter{Symmetry and Support}
In order to analyse families of symmetric circuits we will need to develop some
theory for analysing symmetries. In this section we define the terms needed
(e.g. the notion of a support) and prove basic results. We then prove the first
important theorem in this paper, the support theorem, using these tools. It is
worth noting that we prove the support theorem for arbitrary symmetric circuits
with matrix-symmetric gates, not just for rank gates.

\section{Supports and Supporting Partitions}
In this section we formally define a support and the related notion of a
supporting partition.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and let $S \subseteq [n]$. Then $S$ is a
  \emph{support} for $G$ if $\stab(S) \subset G$.
\end{definition}

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and $\mathcal{P}$ be a partition of $[n]$.
  Then $\mathcal{P}$ is a \emph{supporting partition} for $G$ if
  $\stab(\mathcal{P}) \subseteq [n]$.
\end{definition}

Notice that if $\mathcal{P}$ is a supporting partition for $G$ and $P \in
\mathcal{P}$ then $\stab([n] \setminus P) \subseteq \stab(\mathcal{P}) \subset
G$, i.e. $[n] \setminus P$ is a support for $G$.

These supporting partitions admits a natural partial order in terms of
coarseness. If $\mathcal{P}, \mathcal{P}'$ we say that $\mathcal{P}'$ is as
course as $\mathcal{P}$ (denoted by $\mathcal{P}' \supseteq \mathcal{P}$) if for
all $x \in \mathcal{P}$ there exists $y \in \mathcal{P}'$ such that $x \subseteq
y$ \cite{AndersonD17}.

Anderson and Dawar \cite{AndersonD17} show that for any permutation group $G
\subseteq \sym_n$ a unique coarsest partition can always be found. We call this
partition the \emph{canonical supporting partition}, and denote it by $\SP (G)$.
This canonical supporting partition, along with the above observation about
supports, will allow us to construct for each group a canonical support.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$. Let $\| SP(G) \| = \min \{\vert [n]
  \setminus P \vert : P \in SP(G) \}$. We say that $G$ has small support if $\|
  SP(G) \| < \frac{n}{2}$.
\end{definition}

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ such that $G$ has small support. Let
  $\consp(G) = [n] \setminus P$, where $P$ is the maximal element of $\SP(G)$.
  We call $\consp(G)$ the canonical support of $G$.
\end{definition}

\begin{remark}
  Include a remark about the importance of asserting the size consideration. Use
  the alternating group as an example.
\end{remark}

We state the following two results from Anderson and Dawar \cite{AndersonD17}.

\begin{lem}
  \label{lem:SP_conjugation}
  Let $G \leq \sym_n$ and $\sigma \in \sym_n$ then $\sigma \SP (G) = \SP(\sigma
  G \sigma^{-1})$.
\end{lem}

\begin{lem}
  For any $G \leq \sym_n$ we have that $\stab (\SP (G)) \subseteq G
  \setstab(\SP(G))$.
\end{lem}

\section{Group Action on Supports}
In this paper, distinct the case of Anderson and Dawar \cite{AndersonD17}, we
will need to develop theory and terminology for dealing with group actions (and
supports) on more than just gates in the circuit. In particular, we use this
theory for understanding the actions on the rows and columns of a labelling of a
gate.

\begin{definition}
  Let $G$ be a subgroup of $\sym_n$ and $X$ be a set on which a left group
  action of $G$ on $X$ is defined. We denote the canonical supporting partition
  of $x\in X$ by $\SP (x) = \SP (\stab(x))$. We say that $x \in X$ has small
  support if $\stab(x)$ has small support. We say that $X$ has small supports if
  all $x \in X$, $x$ have small support.

  If $x \in X$ has small support, we denote the \emph{canonical support} of $x$
  by $\consp(x) = \consp (stab(x))$.
\end{definition}

\begin{definition}
  Let $G$ be a group and $X$ be a set on which a left group action of $G$ on $X$
  is defined. Then $\stab_G (x) = \{\pi \in G : \pi x = x\}$ and $\orb_G (x) =
  \{\pi x : \pi \in G\}$. In the event that the context makes the group obvious
  we omit the subscript.
\end{definition}

\begin{lem}
  \label{lem:stab_conjugation}
  Let $G$ be a subgroup of $\sym_n$, $X$ be a set on which a left group action
  of $G$ on $X$ is defined and $\sigma \in G$. Then for any $x \in X$ it follows
  that $\sigma \stab (x) \sigma^{-1} = \stab(\sigma x)$.
\end{lem}

\begin{proof}
  Let $\pi \in \stab(x)$, then $\sigma \pi \sigma^{-1}(\sigma x) = \sigma \pi
  (x) = \sigma (x)$, and so $\sigma \pi \sigma^{-1} \in \stab(\sigma x)$. Let
  $\pi \in \stab(\sigma (x))$ then $\pi \sigma (x) = \sigma (x)$ and so
  $\sigma^{-1} \pi \sigma (x) = x$. It follows that $\sigma^{-1} \pi \sigma \in
  \stab(x)$ and so $\pi = \sigma (\sigma^{-1} \pi \sigma) \sigma ^{-1} \in
  \sigma \stab(x) \sigma^{-1}$.
\end{proof}

\begin{lem}
  \label{lem:support_mapping}
  Let $G $ be a subgroup of $\sym_n$, $X$ be a set on which a left group action
  of $G$ on $X$ is defined and $\sigma \in G$. Then for any $x \in X$ it follows
  that $\sigma \SP (x) = \SP (\sigma x)$ and, if $x$ has small support, $\sigma
  \consp (x) = \consp (\sigma x)$.
\end{lem}
\begin{proof}
  From Lemma \ref{lem:SP_conjugation} and Lemma \ref{lem:stab_conjugation} have
  that $\sigma \SP (x) = \SP(\sigma \stab(h) \sigma^{-1}) = \SP (\stab(\sigma
  x)) = \SP (\sigma x)$. Thus proving the first part of the statement.

  From the fact that $\| \SP(x) \| < \frac{n}{2}$ it follows there exists a
  unique $P_1 \in \SP(x)$ such that $\vert P_1 \vert > \frac{n}{2}$ and
  $\consp(x) = [n] \setminus P_1$. But then $\sigma \consp (x) = \sigma([n]
  \setminus P_1) = [n] \setminus (\sigma P_1)$. We note that $\sigma P_1 \in
  \sigma \SP (x) = \SP(\sigma x)$ and $\vert \sigma P_1 \vert > \frac{n}{2}$.
  Thus $\sigma P_1$ is the unique largest partition in $\SP(\sigma x)$, and so
  $\consp(\sigma x) = [n]\setminus (\sigma P_1) = \sigma (consp(x))$. Thus
  proving the second part of the statement.
\end{proof}

\begin{lem}
  Let $G$ be a subgroup of $\sym_n$, $X$ be a set on which a left group action
  of $G$ on $X$ is defined. Then for any $x \in X$ with small support, and
  $\sigma, \sigma' \in G$ such that $\sigma (vec{\consp}(x)) = \sigma'
  (\vec{\consp}((x))$ we have that $\sigma (x) = \sigma'(x)$.
\end{lem}
\begin{proof}
  We have that $(\sigma')^{-1}\sigma(\vec{\consp}(x)) = \vec{\consp}(x)$ and so
  $(\sigma')^{-1} \sigma (x) = x$ and thus $\sigma (x) = \sigma' (x)$.
\end{proof}

% \begin{lem}
%   Let $G \leq \sym_n$, $X$ be a set on which a left group action of $G$ on $X$
%   is defined and $\sigma \in G$. Then if $y \in \orb_{G} (x)$ and $\consp (x)
%   = \consp (y)$ then $x = y$.
% \end{lem}

% \begin{proof}
%   Let $\sigma \in G$ be such that $\sigma x = y$. Then $\consp(x) = \consp (y)
%   = \consp (\sigma x) = \sigma \consp (x)$. Thus $\sigma (\sp (x))$
% \end{proof}

% \begin{definition}
%   Let $h \in H$. Let $\stab_r(h)$ be a subgroup of $\stab(\consp(g))$
%   consisting of exactly those permutations $\sigma$ such that $\row (\sigma h)
%   = \row (h)$. Let $\stab_c(h)$ be a subgroup of $\stab(\consp(g))$ consisting
%   of exactly those permutations $\sigma$ such that $\column (\sigma h) =
%   \column(h)$.
% \end{definition}

% \begin{definition}
%   A set $S \subseteq [n]$ is called a row-support for $h$ if $S$ is a support
%   for $\stab_r(h)$. A set $S \subseteq [n]$ is called a column-support for $h$
%   iff $S$ is a support for $\stab_c(h)$.
% \end{definition}

>>>>

\documentclass[../paper.tex]{subfiles}
\begin{document}
A finite relational structure, such as a graph, can be represented as a binary
string in a standard way (for instance, by enumerating its adjacency matrix).
When we speak of the computational complexity of properties of graphs, we assume
such an encoding in order to present the structure as input to a machine or
circuit or other appropriate computational model. The properties of structures
we are interested in studying are invariant under isomorphisms. This requirement
translates into an invariance condition on sets of strings---say that a language
$L \subseteq \{0,1\}^\star$ is a \emph{graph property} if, whenever $x \in L$
for a string $x$ representing a graph $G$, we have $y\in L$ for any string $y$
obtained from $x$ by re-ordering the vertices of $G$. The central open question
in descriptive complexity---whether there is a logical characterisation of the
polynomial-time decidable properties of finite relational structures--- can then
be understood as the question of characterizing those languages $L$ that are in
$\PT$ and are graph properties.

@@

This motivates the study of symmetric circuit families that involve gates which
themselves may compute non-symmetric functions. In particular, motivated by the
study of rank logic~\cite{DGHL09,GP15}, we consider gates that are
\emph{matrix-symmetric}. These are gates computing a Boolean function $f:
\{0,1\}^{A \times B} \rightarrow \{0,1\}$ which is invariant under arbitrary
permutations of $A$ and $B$. A typical example is the function computing the
rank of an $A \times B$ $0$-$1$ matrix. We show that the support theorem of
Anderson and Dawar can be shown to hold, even in the presence of
matrix-symmetric gates. This allows us to show that definability in the logic
fixed-point with rank ($\FPR$) is the same as decidability by
polynomially-uniform families of symmetric circuits with rank gates.
% Another extension of $\FPC$ that has received much attention in the
% descriptive complexity literature is the class $\CPTC$ of \emph{choiceless
% polynomial time with counting}. The work of~\cite{AD14} leaves open the
% question of obtaining a circuit characterization of this class and it was
% suggested there that this would require a relaxation of the notion of symmetry
% used in defining symmetric circuits. In particular, while $\CPTC$ is based on
% a machine model designed to preserve symmetry by forbidding choices, the
% polynomial time and space restrictions apply int he context of the symmetries
% fo a \emph{single input structure}. In this paper, we consider a relaxation of
% the notion of symmetric circuit intended to capture exactly this intuition.
% However, we are able to show that this does not yield a characterization of
% $\CPTC$. In particular, we show that no family in this class of relaxed
% symmetric circuits can distinguish odd from even Cai-F\"urer-Immerman graphs,
% something we know is do-able in $\CPTC$~\cite{DRR08}.

\end{document}
